Learning by Teaching a Robot:\\The Case of Handwriting

### 2015-12-11 main.tex
Imported some bits (including Normandy study) from HRI paper

\usepackage{graphicx} 
\usepackage{subfigure}
\usepackage{paralist}

\usepackage{hyperref}

@@

\author{S\'everin Lemaignan$^1$, Alexis Jacq$^{1,2}$, Fernando Garcia$^1$,
    Deanna Hood$^1$, \\Aude
    Billard$^1$, Ana Paiva$^2$, Pierre Dillenbourg$^1$ \\
$^1$CHILI Lab, \'Ecole Polytechnique F\'ed\'erale de Lausanne, Suisse,\\
$^2$Instituto Superior T\'{e}cnico, University of Lisbon, Portugal}

@@

Besides the commitment of the child into the interaction build on the
``prot\'eg\'e effect'': the teacher feels responsible for his student, commits
to the student's success and possibly experiences student's failure as his own
failure to teach. Teachable computer-based agents have previously been used to
encourage this ``prot\'eg\'e effect'', wherein students invest more effort into
learning when it is for a teachable agent than for themselves~\cite{Chase2009}.
We rely on this cognitive mechanism to reinforce the child's commitment into the
robot-mediated handwriting activity.

@@

In our case, the role of agency is stronger: it induces meta-cognition (``I am
interacting with an agent, so I need to reflect on how to best teach him'')
which is beneficial for the learning process; it also induces a prot\'eg\'e
effect (``I want my robot-agent to succeed!'') which support the commitment of
child into the interaction, also for longer period of time (we report in this
article upon two case studies where children were indeed interacting with the
same robot for several hours, spanned over several weeks).

@@

The last session was set as a test: the ``explorer'' robot
had come back from its mission and it actually challenged the other robot in
front of Vincent: \emph{``I don't believe you wrote yourself these nice letters that I
received! Prove it to me by writing something in front of me!''} This situation
was meant to evidence the Prot\'eg\'e effect: by judging the other robot's
handwriting, the ``explorer'' robot would implicitly judge Vincent's skills as
teacher, and in turn, Vincent's handwriting.

@@

Overall, Vincent provided 154 demonstrations to the robot, and he remained
actively engaged over the four weeks. The story was well accepted by Vincent and
he seriously engaged into the game. After the first week, he showed good
confidence to play with the robot and he built affective bonds with the robot
over the course of the study, as evidenced by some cries on the last session,
and a letter sent by him to the robot 4 months \emph{after} the end of the study
to get news. This represents a promising initial result: we can effectively keep
a child engaged with the robot for a relatively long period of time (about 5
hours).

No conclusion can be drawn in terms of actual handwriting remediation: we did
not design this study to formally assess possible improvements.  However, as
pictured on Figure~\ref{fig:stimuli}, Vincent was able to significantly improve
the robot's skill, and he acknowledged that he had been able to help the robot:
in that regard, Vincent convinced himself that he was ``good enough'' at writing
to help someone else, and this is likely to have positively impacted his
self-esteem.

@@

The experiment took place at the therapist's surgery (four sessions spanning
over 5 weeks). Contrary to Vincent's experiment, we chose not to introduce any
storyline beyond a simple prompt (``the robot wants to participate to a robotic
handwriting contest, will you help him to prepare?''), only provided during the
first session. Hence, we also tested during this case-study if the robot (and
the Prot\'eg\'e effect) would induce a strong enough intrinsic motivation to
keep the child engage over the five weeks.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{learning_6_demos}
    \caption{\small Thomas' demonstrations for the number ``6'' (top row) and
        the corresponding shapes generated by the robot (bottom row). After
        eight demonstrations, Thomas decided that the robot's ``6'' was good enough,
        and went to another character. In that regard, he was the one
        responsible for the learning process of the robot.}

    \label{learning_6_demos}
\end{figure}

@@

Despite his attention deficit, Thomas was able to conduct the activity during
more than forty minutes in each session. In total, 55 allographs out of the 82
shapes demonstrated by the child were accepted by the robot (with a progressive
improvement from 13 out of 28 in the first session up to 26 out of 29 in the
last session).

As soon as Thomas understood that the robot was only accepting well-formed
allographs, he started to focus on it and he would typically draw 5 or 6 times
the number before actually sending to the robot (the tablet let the children
clear their drawing and try again before sending it to the robot). According to
the therapist, it was the first time that Thomas would correct himself in such a
way, explicitly having to reflect on how \emph{another agent} (the robot) would
interpret and understand his writing.

Since the robot's handwriting started from a simple primitive (a stroke), each
time Thomas succeeded to have his demonstration accepted by it, the robot's
improvement was clearly visible (as measured in Figure~\ref{henry_distances}).
This led to a self-rewarding situation that effectively supported Thomas'
engagement.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{learning_6_distances}
    \caption{\small \textbf{Assessing handwriting progress}: Euclidian
    distance in the eigenspace of the number ``6''. Green lines represent the robot performance,
    blue lines Thomas's performance. The round IDs ($x$-axis) correspond to the demonstrations
pictured on Figure~\ref{learning_6_demos}.}
    \label{henry_distances}
\end{figure}

@@

\begin{itemize}
    \item robots in an educative context are \textbf{relevant and effective beyond ICT
        teaching},

>>>>

\usepackage[utf8]{inputenc}

\usepackage{graphicx} 
\usepackage{subfigure}
\usepackage{paralist}
\usepackage{amsfonts}
\usepackage{hyperref}

@@

\author{Séverin Lemaignan$^1$, Alexis Jacq$^{1,2}$, Fernando Garcia$^1$,
    Deanna Hood$^1$, \\Aude
    Billard$^1$, Ana Paiva$^2$, Pierre Dillenbourg$^1$ \\
$^1$CHILI Lab, École Polytechnique Fédérale de Lausanne, Suisse,\\
$^2$Instituto Superior Técnico, University of Lisbon, Portugal}

@@

Besides the commitment of the child into the interaction build on the
``protégé effect'': the teacher feels responsible for his student, commits
to the student's success and possibly experiences student's failure as his own
failure to teach. Teachable computer-based agents have previously been used to
encourage this ``protégé effect'', wherein students invest more effort into
learning when it is for a teachable agent than for themselves~\cite{Chase2009}.
We rely on this cognitive mechanism to reinforce the child's commitment into the
robot-mediated handwriting activity.

@@

In our case, the role of agency is stronger: it induces meta-cognition (``I am
interacting with an agent, so I need to reflect on how to best teach him'')
which is beneficial for the learning process; it also induces a protégé
effect (``I want my robot-agent to succeed!'') which support the commitment of
child into the interaction, also for longer period of time (we report in this
article upon two case studies where children were indeed interacting with the
same robot for several hours, spanned over several weeks).

@@

The last session was set as a test: the ``explorer'' robot
had come back from its mission and it actually challenged the other robot in
front of Vincent: \emph{``I don't believe you wrote yourself these nice letters that I
received! Prove it to me by writing something in front of me!''} This situation
was meant to evidence the Protégé effect: by judging the other robot's
handwriting, the ``explorer'' robot would implicitly judge Vincent's skills as
teacher, and in turn, Vincent's handwriting.

@@

Overall, Vincent provided 154 demonstrations to the robot, and he remained
actively engaged over the four weeks. The story was well accepted by Vincent and
he seriously engaged into the game. After the first week, he showed good
confidence to play with the robot and he built affective bonds with the robot
over the course of the study, as evidenced by some cries on the last session,
and several letters sent to the robot \emph{after} the end of the study
(one of them 4 months later) to get news. This represents a promising initial
result: we can effectively keep a child engaged with the robot for a relatively
long periods of time (about 5 hours).

No conclusion can be drawn in terms of actual handwriting remediation: we did
not design this study to formally assess possible improvements. However, as pictured on Figure~\ref{fig:stimuli}, Vincent was able to
significantly improve the robot's skill, and he acknowledged that he had been
able to help the robot: in that regard, Vincent convinced himself that he was
``good enough'' at writing to help someone else, which is likely to have
a positive impact on his self-esteem.

@@

The experiment took place at the therapist's surgery (four sessions spanning
over 5 weeks). Contrary to Vincent's experiment, we chose not to introduce any
storyline beyond a simple prompt (``the robot wants to participate to a robotic
handwriting contest, will you help him to prepare?''), only provided during the
first session. Hence, we also tested during this case-study if the robot (and
the Protégé effect) would induce a strong enough intrinsic motivation to
keep the child engage over the five weeks.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{learning_6_demos}
    \caption{\small Demonstrations provided by Thomas for the number ``6'' (top row) and
        corresponding shapes generated by the robot. After eight demonstrations,
        Thomas decided that the robot's ``6'' was good enough, and went to
    another character: in that respect, he was the one leading the learning
process of the robot.}
    \label{learning_6_demos}
\end{figure}

@@

Despite his attention deficit, Thomas was able to remain engaged in the activity during more than
forty minutes in each session. In total, 55 allographs out of 82 
demonstrated by the child were acceptable considering our threshold (with a
progressive improvement from 13 out of 28 in the first session up to 26 out
of 29 in the last session).

As soon as Thomas understood that the robot was only accepting well-formed
allographs, he started to focus on it and he would typically draw 5 or 6 times
the number before actually sending to the robot (the tablet lets children
clear their drawing and try again before sending it). According to
the therapist, it was the first time that Thomas would correct himself in such a
way, explicitly having to reflect on how \emph{another agent} (the robot) would
interpret and understand his writing. Figure~\ref{Thomas_progress} shows how
he gradually improved his demonstrations for some numbers, according to the
metric we used to make the robot accept/refuse samples.

Since the robot's handwriting started from a simple primitive (a stroke), each
time Thomas succeeded to have his demonstration accepted by it, the robot's
improvement was clearly visible (as measured in Figure~\ref{Thomas_distances}).
This led to a self-rewarding situation that effectively supported Thomas'
engagement.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{learning_6_distances}
    \caption{\small Two metrics to assess the handwriting progresses: Euclidean
    distance in the subspace of the number dataset (top figure) or in
Cartesian space (bottom figure). Green lines represent the robot performance,
blue lines performance of the child. The round IDs correspond to the demonstrations
pictured on Figure~\ref{Thomas_demos}.}
    \label{Thomas_distances}
\end{figure}

\begin{figure}
    \centering
    \subfigure[number 1]{
        \includegraphics[width=6cm]{henry1}
    }
    \subfigure[number 2]{
        \includegraphics[width=6cm]{henry2}
    }
	\subfigure[number 5]{
        \includegraphics[width=6cm]{henry5}
    }
    \caption{\small Improvement of Thomas demonstrations for some numbers: a) drawing the allograph of the number 1, b) the number 2 and c) the number 5. Thomas progressively took care of the demonstrations he was providing to the robot for those numbers. We used for this figure the same metric than the one used for the acceptance algorithm. Distances are normalized with respect to the biggest value. The dashed line correspond to the threshold of robot's acceptance.}

    \label{Thomas_progress}
\end{figure}

\section{Case study 3: when children evaluate the robot}\label{auto}

\subsection{Context}

Each of previous studies was adapted to individual subjects : we used a special
design for each case in order to sustain the child's engagement.
In our next experiment, we conducted case studies with eight children using a single
design. Those children have in common difficulties to learn
cursive writing but the natures and intensities of those troubles are radically
different from one child to another. Valentine (7 years old), Alexandre (6.5) and
Jonathan (7) are under the care of an
occupational therapist. Enzo (8) and Matenzo (7) are repeating their school year
because of writing. Mona (6) and Adele (8) are bottom of their respective
classes in writing activities. Nathan (7) is under the care of a neurologist, and
has been diagnosed with specific language impairment. All of those children are
expected, given their school level, to know the allographs of
cursive letters.

\subsection{Questions}

The main purpose of this study is to test the ergonomic aspects of CoWriter. When introducing the robot, we did not
provide scenario and gave to children minimal explanations. We then observed how easily children assumed the role of the teacher and how seriously they tried to help the robot.

\subsection{Experimental settings}

This experiment took place in the coffee room of a therapists shared surgery
in Normandy, France. Over two weeks, each child came three time for one-hour
session, except Adele and Mona who did just one session. A facilitator was present to explain the rules of the game and tablet usage. As in the previous experiment, children were provided with two tablets : one to choose a word (or a single letter) to teach, and one
used by both the child and the robot to write. We also provided the allograph's template if the child asked for.

The starting point of the robot's writing was the same for all children: we
used middle point between simple vertical strokes and letters. For this study,
we wanted the robot to be only influenced by the demonstrations provided by the
child, so we did not project allographs in a subspace. The new generated
sample by the robot were calculated as the middle way between demonstration and
last state in Cartesian space.

The robot was programmed to accept all demonstrations, giving to the child the
full responsibility of the teacher.

We added two buttons on the tablet  interface: a green one with a ``thumbs up", and
a red one with a ``thumbs down". Those buttons could be used by children to evaluate the
robot (the green one was for rewards while the red one was for punishment). This
way, we could measure the perception of the robot by the child: the more the
child used evaluation buttons, the more he was playing the teacher, judging the
robot instead of himself. It became possible to estimate if a child is playing
seriously given correlation between his evaluation and robot's actual
progression.

\subsection{Results}

All children maintained their engagement during the whole sessions. They
provided on average 42 demonstrations per session. All children used evaluation
buttons and had preference to reward the robot (in total, 99 rewards and 33
punishments were recorded).

Since sessions took place over only two weeks, we did not studied possible
handwriting remediation in children, but we focused on correlation between the
children's evaluations and the robot's progression.  We estimated the robot's
progression as the difference between a starting score (score of the first
robot's try when children have chosen a new word/letter to work on) and the
current robot's score (after being taught by the child). The score is calculated
as the average of euclidean distance between robot's try and a reference
allograph over all letters of the word. Those references for letter allographs
where drawn by us beforehand, taking inspiration in education.com cursive
letters
template\footnote{\url{http://www.education.com/slideshow/cursive-handwriting-z/}}.
Let $P_i$ be the estimated progression of the robot at time $ i$. Of course, if
the child chose to switch to a new word at time $ j$ , we have $ P_{j}=0$ and
clearly $ P_0=0$.  To measure whether the fact that the child was rewarding the
robot when it was progressing was significant or not, we generated 10000 times
the same number of reward/punishment but accorded at random times. Let $ R_i^n$
be the $n^{th}$ generated evaluation at time $ i$ ($ R_i^n = 0$ if no evaluation
occurred at time $ i$ , $ R_i^n=1$ if a reward occurred at time $i$ and
$R_i^n=-1$ if a punishment occurred at time $i$), and $\overline{R}_i$ be the
actual evaluation at time $i$. For each $n^{th}$ generated sequence of
evaluation, we compute a score of evaluation: $$ S^n = \sum\limits_i{R_i^n
P_i}$$ Then we can estimate the p-value $p$ of the actual score: $$ \overline{S}
= \sum\limits_i{\overline{R}_i P_i}$$ given the distribution of the generated
scores $\left(S^n\right)_n$, which is assumed to be Gaussian:

$$
p(\overline{S})= \mathbb{P}{\left[X>\overline{S}\right]} = 1-\phi{\left(\frac{\overline{S}-\mu}{\sigma}\right)}
$$

where $\phi$ denotes the cumulative distribution function of the standard normal
distribution, $\mu$ the mean and $\sigma$ the deviation of the generated scores
$\left(S^n\right)_n$.  As a result, we found that 5 of the 8 children obtained a
score of evaluation significantly hight ($p(\overline{S})<0.05$). Score of
evaluation p-values of each child is reported in the second-to-last column of
Table~\ref{table:scores}.

We also studied correlation between children's evaluations and their own
progression. The analysis was conducted in the same way, using distances between
children demonstrations and reference allographs to compute children
progressions.  3 of 5 children that played ``seriously" obtained score of
evaluation of their own progression significantly high (last column of
Table~\ref{table:scores}). For those children, it seems that the robot was
reflecting their own performances, and while they were judging the robot
positively (three times more rewards than punishments) they were actually
evaluating themselves.

\begin{table}
    \centering
    \begin{tabular}{|c|c|c|c|c|c|}
        \hline
        Child & demo & rew & pun & p(robot) & p(child)\\ \hline
        Valentine & 127 & 24 & 6 & 2.4e-03 & 5.5e-02\\ \hline
        Enzo & 223 & 20 & 9 & 1.7e-01 & 3.5e-01\\ \hline
        Matenzo & 131 & 10 & 3 & 3.8e-03 & 7.9e-03\\ \hline
        Jonathan & 98 & 10 & 5 &  1.5e-01 & 3.8e-01\\ \hline
        Nathan & 115 & 16 & 4 & 5.3e-04 & 2.7e-03\\ \hline
        Alexandre & 83 & 10 & 3 & 3.1e-02 & 6.0e-01\\ \hline
        Adele & 35 & 4 & 2 & 5.0e-02 & 3.7e-02\\ \hline
        Mona & 40 & 5 & 1 &  5.4e-01 & 2.0e-01\\ \hline
    \end{tabular}
    \caption{\small Results of evaluations. demo: number of demonstrations provided by
        the child over all session. rew: number of rewards accorded by
        the child. pun: number of punishments. p(robot): are the evaluations
        significantly corresponding to the progression of the robot ? p(child): are the
        evaluations significantly corresponding to the child's own
progression ?}
    \label{table:scores}
\end{table}

@@

\begin{itemize}
    \item robots in an educative context are \textbf{relevant and effective
        beyond STEM} (Science, Technology, Engineering and Mathematics)
    \textbf{teaching},
